---
title: "MACROM: Optimal Control Workflow for Climate Temperature Overshoot"
author: "Nina Rynne"
institution: "Griffith University"
output: html_document
date: "January 2026"
---

# ==============================================================================
# MACROM: An Optimal Control Model for Balancing Climate Change 
# Abatement and Damage Trade-offs
# 
# Authors: Nina Rynne, Michael Bode, Melanie Roberts, Ryan Heneghan
# Institution: Griffith University
# 
# Copyright (c) 2025 Nina Rynne
# Licensed under CC-BY-4.0 - see LICENSE file for details
# 
# Citation: If you use this code, please cite:
#   Rynne, N., Bode, M., Roberts, M., & Heneghan, R. (2025). 
#   MACROM: An Optimal Control Model for Balancing Climate Change 
#   Abatement and Damage Trade-offs.
#   https://doi.org/10.5281/zenodo.18463951
# 
# Version: 2.0.0
# Last updated: March 2026
# 
# Repository: https://github.com/nina-rynne/MACROM
# Contact: nina.rynne@griffithuni.edu.au
# ==============================================================================

# ABOUT MACROM:
# This workflow runs the complete MACROM optimal control model. MACROM determines
# optimal climate mitigation and CDR strategies by balancing abatement costs 
# against temperature-related damages under different SSP emission scenarios.
#
# This is the primary user interface - public users interact only with this file.
# For detailed instructions, prerequisites, and troubleshooting, see USER_GUIDE.md

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Create output directories if they don't exist
dir.create(here::here("output"), showWarnings = FALSE)
dir.create(here::here("figs"), showWarnings = FALSE)

```

```{r libraries}

# ==============================================================================
# Required Libraries
# ==============================================================================
# Install any missing packages using: install.packages("package_name")

# ============================================
# Data Manipulation and Transformation
# ============================================
library(purrr)        # Functional programming tools
library(tidyverse)    # Data science toolkit

# ============================================
# File Handling and I/O
# ============================================
library(readr)        # Reading data files
library(here)         # File path management
library(yaml)         # YAML file handling
library(Cairo)        # PDF export

# ============================================
# Parallel Processing
# ============================================
library(parallel)     # Parallel processing
library(foreach)      # Parallel iteration
library(doParallel)   # Parallel backend

# ============================================
# Data Analysis and Statistics
# ============================================
library(lhs)          # Latin hypercube sampling
library(iterators)    # Iterator support

# ============================================
# Visualisation
# ============================================
library(ggplot2)      # Data visualisation
library(cowplot)      # Plot arrangement
library(patchwork)    # Plot composition
library(viridisLite)  # Colourblind friendly palettes
library(viridis)      # Colourblind friendly palettes
library(colorspace)   # Advanced colour manipulation
library(scales)       # Scale functions for visualisation
library(metR)         # Contour line labelling
library(gridExtra)    # Plot arrangement
```

```{r data_preparation_call}
# ==============================================================================
# Data Preparation
# ==============================================================================
# Import and prepare emissions and economic data
# This section imports raw SSP scenario data and interpolates it to annual time steps

# Source data preparation functions
source(here::here("src", "data_preparation.R"))

# Import emissions and economic data
emissions_imported <- import_ssp_emissions("emissions.csv")
economic_imported <- import_ssp_economic("gwp.csv")

# Interpolate data to annual time step (2020-2100)
# USER CHOICE: Modify start_year, end_year, or dt if different time range/step is needed
emissions_df <- interpolate_ssp_emissions(
  emissions_df = emissions_imported,
  dt = 1,
  start_year = 2025,
  end_year = 2100
)

economic_df <- interpolate_ssp_economic(
  economic_df = economic_imported,
  dt = 1,
  start_year = 2025,
  end_year = 2100
)
```

```{r data_preparation_2500_call}
# ==============================================================================
# Data Preparation — Extended Horizon (2020-2500)
# ==============================================================================
# Import and prepare emissions and economic data extended to 2500 on a
# decadal time step (data source: emissions_2500.csv, gwp_2500.csv).
# Same import/interpolation logic as data_preparation_call above, just with
# the longer-horizon source files. Used by analyses that need a genuine
# (non flat-lined) horizon beyond 2100, e.g. free_time_scale_sensitivity_analysis.

# Source data preparation functions
source(here::here("src", "data_preparation.R"))

# Import emissions and economic data (extended to 2500)
emissions_imported_2500 <- import_ssp_emissions("emissions_2500.csv")
economic_imported_2500  <- import_ssp_economic("gwp_2500.csv")

# Interpolate data to annual time step (2020-2500)
# USER CHOICE: Modify start_year, end_year, or dt if different time range/step is needed
emissions_df_2500 <- interpolate_ssp_emissions(
  emissions_df = emissions_imported_2500,
  dt = 1,
  start_year = 2025,
  end_year = 2500
)

economic_df_2500 <- interpolate_ssp_economic(
  economic_df = economic_imported_2500,
  dt = 1,
  start_year = 2025,
  end_year = 2500
)
```

```{r latin_hypercube_sampling_call, eval=FALSE}
# ==============================================================================
# PARAMETER APPROACH 1: Multiple Parameter Sets (Sensitivity Analysis)
# ==============================================================================
# ⚠️ USER CHOICE: Use EITHER this chunk OR the fixed_model_parameters_call chunk below
# Do NOT run both - they are mutually exclusive options
#
# This approach generates multiple parameter combinations for uncertainty quantification
# Parameter ranges defined in: parameter_details.yml
#
# WHEN TO USE:
# - Exploring parameter sensitivity
# - Uncertainty quantification
# - Generating confidence intervals
#
# NOTE: Significantly increases computation time
# Recommended: 100-500 samples for thorough analysis

source(here::here("src", "latin_hypercube_sampling.R"))
source(here::here("src", "model_parameters.R"))

# USER CHOICE: Change to desired number of parameter sets to generate
# More samples = better coverage but longer runtime
lhs_parameter_df <- generate_lhs_samples(n_samples = 100, seed = 12345)

# Add fixed model parameters to the LHS samples
parameter_df <- add_fixed_parameters(lhs_parameter_df)
```

```{r fixed_model_parameters_call}
# ==============================================================================
# PARAMETER APPROACH 2: Single Parameter Set (Recommended for most analyses)
# ==============================================================================
# ⚠️ USER CHOICE: Use EITHER this chunk OR the latin_hypercube_sampling_call chunk above
# Do NOT run both - they are mutually exclusive options
#
# This approach uses a single predefined parameter set
# Parameter values defined in: parameter_details_fixed.yml
#
# WHEN TO USE:
# - Comparing across SSP scenarios
# - Analysing delayed deployment effects
# - Generating primary results with best-estimate parameters
# - Quick test runs
#
# RECOMMENDED FOR: Most standard analyses and publication figures

source(here::here("src", "model_parameters.R"))

# Use predefined parameter set
parameter_df <- create_params_dataframe()
```

```{r scenario_comparison_analysis, eval=FALSE}
# ==============================================================================
# Scenario Comparison Analysis
# ==============================================================================
# Compare optimal control solutions across different SSP baseline scenarios
# 
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: ~5-15 minutes for 5 scenarios with parallel processing

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "scenario_comparison.R"))

# USER CHOICE: Select which SSP scenarios to compare
scenarios_to_compare <- c("SSP5-Baseline")

# USER CHOICE: Delay settings (typically keep at 0 for baseline comparison)
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay <- 0         # Years before CDR deployment begins

# Run scenario comparison
scenario_comparison_results <- run_scenario_comparison(
  parameter_df = parameter_df[1, ],      # Must be single row
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenarios = scenarios_to_compare,
  mitigation_delay_years = mitigation_delay,
  cdr_delay_years = cdr_delay,
  use_parallel = TRUE,                   # Set FALSE if parallel issues occur
  save_results = TRUE,                   # Automatically saves to output/
  verbose = FALSE                        # Set TRUE for detailed progress
)

# Display summary
if (!is.null(scenario_comparison_results$comparison_summary)) {
  print(scenario_comparison_results$comparison_summary)
}
```

```{r scenario_comparison_visualisation, eval=FALSE}
# ==============================================================================
# Scenario Comparison Visualisation
# ==============================================================================
# Create publication-quality visualisations comparing optimal control solutions
# OUTPUTS: Saved automatically to figs/ directory

# Source visualisation functions
source(here::here("src", "scenario_comparison_visualisation.R"))

# Check if results exist
if (!exists("scenario_comparison_results")) {
  stop("Scenario results not found. Please run the scenario_comparison_analysis chunk first.")
}

# Create comprehensive dashboard
dashboard <- create_scenario_comparison_dashboard(
  scenario_results = scenario_comparison_results,
  
  # USER CHOICE: Customise dashboard
  include_temperature = TRUE,
  include_emissions = TRUE,
  include_controls = TRUE,
  include_costs = TRUE,
  
  # USER CHOICE: Styling
  color_palette = "viridis",
  save_plot = TRUE,
  width = 297,      # A4 landscape width (mm)
  height = 210,     # A4 landscape height (mm)
  verbose = TRUE
)

# Display the dashboard
print(dashboard)
```

```{r parameter_importance_analysis, eval=FALSE}
# ==============================================================================
# Solution generation for multiple parameters & single SSP
# ==============================================================================
# Parameter importance analysis using Latin Hypercube Sampling (LHS) parameter sets
# This explores the solution space across multiple parameter combinations
# Provides output for sensitivity analysis

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "parameter_importance.R"))

# USER CHOICE: Modify these settings as needed
scenario_to_analyse <- "SSP4-Baseline"

# Run parameter importance analysis (saves automatically)
parameter_importance_results <- run_parameter_importance(
  parameter_df = parameter_df,
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenario = scenario_to_analyse,
  mitigation_delay_years = 0,
  cdr_delay_years = 0,
  use_parallel = TRUE,
  save_results = TRUE,
  verbose = FALSE
)

# Filter parameter importance results to remove problematic runs
# This removes runs that fail to return temperature to 1.5°C by 2100

# USER CHOICE: Set maximum acceptable final temperature (default: 1.6°C)
max_acceptable_temperature <- 1.6

# Apply filtering
parameter_importance_results_cleaned <- filter_importance_results(
  importance_results = parameter_importance_results,
  max_final_temperature = max_acceptable_temperature,
  verbose = TRUE
)
```

```{r parameter_importance_visualisation, eval=FALSE}
# ==============================================================================
# Plot multiple parameters & single SSP results
# ==============================================================================
# Create visualisations for parameter importance analysis results

# Source parameter importance visualisation functions
source(here::here("src", "parameter_importance_visualisation.R"))

# Check if results exist
if (!exists("parameter_importance_results_cleaned")) {
  stop("Parameter importance results not found. Please run the parameter_importance_analysis and parameter_importance_filtering chunks first.")
}

# Create and display dashboard with insights
parameter_importance_dashboard <- create_parameter_importance_dashboard(
  importance_results = parameter_importance_results_cleaned,
  save_plot = TRUE,
  print_insights = FALSE,
  verbose = FALSE
)

print(parameter_importance_dashboard)
```

```{r export_parameter_importance_for_sobol, eval=FALSE}
# ==============================================================================
# Export multiple SSPs & single parameter set results
# ==============================================================================
# Export parameter importance results for external Sobol indices analysis
# Creates CSV files with parameter values and output variables

# Source required functions
source(here::here("src", "parameter_importance.R"))

# Check if cleaned results exist
if (!exists("parameter_importance_results_cleaned")) {
  stop("Cleaned parameter importance results not found. Please run parameter_importance_filtering chunk first.")
}

# Export results for Sobol analysis (saves automatically)
export_for_sobol_analysis(
  importance_results = parameter_importance_results_cleaned,
  verbose = FALSE
)
```

```{r delayed_deployment_analysis, eval=FALSE}
# ==============================================================================
# Delayed Deployment Analysis
# ==============================================================================
# Analyse the effects of delaying mitigation and/or CDR deployment
# 
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: Scales with (max_delay / step_size)² × number of scenarios
#          Example: 50-year max, 10-year steps, 5 scenarios = ~30 minutes

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "delayed_deployment.R"))

# USER CHOICE: Delay range and granularity
max_delay_years <- 70        # Maximum delay to test (years)
delay_step_size <- 10          # Step size (1 = yearly, 10 = every decade)

# USER CHOICE: Scenarios to analyse
# Options: "all" or vector like c("SSP2-Baseline", "SSP3-Baseline")
scenarios_to_analyse <- "all"

# Run delayed deployment analysis
delayed_deployment_results <- run_unified_delayed_deployment(
  parameter_df = parameter_df[1, ],      # Must be single row
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenarios = scenarios_to_analyse,
  max_delay_years = max_delay_years,
  delay_step_size = delay_step_size,
  use_parallel = TRUE,
  save_results = TRUE,
  verbose = FALSE
)

# Display summary
cat("\n=== DELAYED DEPLOYMENT ANALYSIS COMPLETE ===\n")
cat("Delay range tested: 0 -", max_delay_years, "years\n")
cat("Total runs:", length(delayed_deployment_results$results), "\n")
```

```{r delayed_deployment_visualisation, eval=FALSE}
# ==============================================================================
# Delayed Deployment Visualisation
# ==============================================================================
# Create heatmap visualisations showing how deployment delays affect outcomes
# OUTPUTS: Saved automatically to figs/ directory

# Source visualisation functions
source(here::here("src", "delayed_deployment_visualisation.R"))

# Check if results exist
if (!exists("delayed_deployment_results")) {
  stop("Delayed deployment results not found. Please run the delayed_deployment_analysis chunk first.")
}

# Create abatement cost field (mitig cost + remov_cost)
# Calculate abatement cost
delayed_deployment_results$combined_results <- delayed_deployment_results$combined_results %>%
  mutate(abatement_cost = mitig_cost + remov_cost)

# Create main dashboard
dashboard <- create_delayed_deployment_dashboard(
  deployment_results = delayed_deployment_results,
  
  # USER CHOICE: Variables to plot
  # Options: "peak_temperature", "years_above_1p5", "abatement_cost", 
  #          "temp_cost", "total_cost", "mitig_cost", "remov_cost"
  variables = c("peak_temperature", "years_above_1p5"),
  
  # USER CHOICE: Visual features
  add_contours = TRUE,
  contour_alpha = 0.6,
  add_arrows = FALSE,
  show_infeasible = FALSE,
  
  # USER CHOICE: Styling
  color_palettes = NULL,
  shared_scale = FALSE,
  # Enable percentile capping
  use_scale_limits = TRUE,
  scale_limit_percentile = 70,
  save_plot = TRUE,
  width = 297,
  height = 210,
  verbose = FALSE
)

# Display the dashboard
print(dashboard)
```

```{r delayed_deployment_prebuilt_visualisation, eval=FALSE}
# ==============================================================================
# Convenience Wrapper Functions for Common Visualisations
# ==============================================================================
# These provide pre-configured shortcuts for standard analysis types

# Temperature and Overshoot Analysis
temperature_dashboard <- create_temperature_delay_dashboard(
  deployment_results = delayed_deployment_results,
  save_plot = FALSE,
  filename = NULL
)
print(temperature_dashboard)

# Cost Analysis with Gradient Arrows
cost_dashboard <- create_cost_delay_dashboard(
  deployment_results = delayed_deployment_results,
  cost_variable = "abatement_cost",  # Options: "abatement_cost", "temp_damage_cost", "total_cost"
  arrow_scale = 3.0,
  min_magnitude = 0,
  save_plot = FALSE,
  filename = NULL
)
print(cost_dashboard)

# Comprehensive Cost Comparison
comprehensive_cost_dashboard <- create_comprehensive_cost_delay_dashboard(
  deployment_results = delayed_deployment_results,
  shared_scale = TRUE,
  save_plot = FALSE,
  filename = NULL,
  height = 260
)
print(comprehensive_cost_dashboard)
```



```{r cdr_scale_sensitivity_analysis, eval=FALSE}
# ==============================================================================
# CDR Scale Sensitivity Analysis
# ==============================================================================
# Sweeps CDR logistic parameters K (carrying capacity, GtCO2/year) and r
# (growth rate) across a grid. Mitigation is zeroed out so CDR is the sole
# control, isolating the effect of CDR scale on climate outcomes.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: Scales with n_K × n_r × n_scenarios (e.g. a 51 × 51 grid across
#          5 scenarios ≈ 30-90 minutes depending on hardware)

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))
source(here::here("src", "cdr_scale_sensitivity.R"))

# USER CHOICE: CDR K range (carrying capacity, GtCO2/year)
K_min <- 25     # Minimum CDR carrying capacity
K_max <- 200    # Maximum CDR carrying capacity
n_K   <- 31     # Number of K values to test

# USER CHOICE: CDR r range (growth rate)
r_min       <- 0.02    # Minimum CDR growth rate
r_max       <- 0.20    # Maximum CDR growth rate
n_r         <- 31      # Number of r values to test
r_log_scale <- FALSE   # Set TRUE for log-spaced r values

# USER CHOICE: Fixed CDR starting conditions
g_initial <- 2      # Starting CDR deployment level (GtCO2/year)
t_start   <- 2025   # Year CDR deployment begins

# USER CHOICE: Scenarios to analyse
scenarios_to_compare <- c("SSP1-Baseline")

cdr_scale_sensitivity_results <- run_cdr_scale_sensitivity_from_params(
  K_min        = K_min,
  K_max        = K_max,
  n_K          = n_K,
  r_min        = r_min,
  r_max        = r_max,
  n_r          = n_r,
  r_log_scale  = r_log_scale,
  g_initial    = g_initial,
  t_start      = t_start,
  scenarios    = scenarios_to_compare,
  parameter_df = parameter_df[1, ],
  emissions_df = emissions_df_2300,
  economic_df  = economic_df_2300
)
```

```{r cdr_scale_sensitivity_delay_analysis, eval=FALSE}
# ==============================================================================
# CDR Scale Sensitivity — Multi-Delay Analysis
# ==============================================================================
# Runs the full K x r sensitivity grid for each of a user-chosen set of CDR
# deployment delays. The result shows how the feasible (K, r) parameter space
# shrinks as the delay before CDR deployment increases.
#
# The logistic capacity curve starts at t_start + delay for each run, so no
# capacity builds during the delay period (represents lack of funding /
# development / political will).
#
# REQUIREMENTS: Must source cdr_scale_sensitivity.R (included below)
# OUTPUTS: Per-delay RDS files + combined RDS saved to output/
# RUNTIME: (n_K × n_r) combinations × n_delays × ~0.27 min/comparison
#   Example: 11 x 11 grid, 4 delays, 1 SSP => 484 comparisons => ~2 min
#   Example: 21 x 21 grid, 4 delays, 5 SSPs => ~5 hrs (run overnight)

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))
source(here::here("src", "cdr_scale_sensitivity.R"))

# USER CHOICE: CDR K range (carrying capacity, GtCO2/year)
K_min <- 25     # Minimum CDR carrying capacity
K_max <- 200    # Maximum CDR carrying capacity
n_K   <- 11     # Number of K values (use 11 for testing, 51 for publication)

# USER CHOICE: CDR r range (growth rate)
r_min       <- 0.02    # Minimum CDR growth rate
r_max       <- 0.20    # Maximum CDR growth rate
n_r         <- 11      # Number of r values (use 11 for testing, 51 for publication)
r_log_scale <- FALSE   # Set TRUE for log-spaced r values

# USER CHOICE: Fixed CDR starting conditions
g_initial <- 2      # Starting CDR deployment level (GtCO2/year)
t_start   <- 2025   # Year CDR deployment nominally begins

# USER CHOICE: Delay values to compare (years)
# These are the delays whose boundaries will be overlaid in visualisation
cdr_delays <- c(0, 20, 40)

# USER CHOICE: Scenarios to analyse
# Use one SSP for testing; all five for publication
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline", "SSP4-Baseline", "SSP5-Baseline")

scale_delay_results <- run_cdr_scale_sensitivity_multi_delay(
  K_min             = K_min,
  K_max             = K_max,
  n_K               = n_K,
  r_min             = r_min,
  r_max             = r_max,
  n_r               = n_r,
  r_log_scale       = r_log_scale,
  g_initial         = g_initial,
  t_start           = t_start,
  cdr_delay_years   = cdr_delays,
  scenarios         = scenarios_to_compare,
  parameter_df      = parameter_df[1, ],
  emissions_df      = emissions_df,
  economic_df       = economic_df,
  use_parallel      = TRUE,
  save_results      = TRUE
)

# Flatten to a single data frame for visualisation
scale_delay_flat <- combine_multi_delay_results(scale_delay_results)
```

```{r cdr_scale_sensitivity_delay_heatmaps, eval=FALSE}
# ==============================================================================
# CDR Scale Sensitivity — Outcome Heatmaps by Delay
# ==============================================================================
# Produces one standard outcome heatmap per CDR delay value, using the same
# recoverable / unrecoverable classification as the base scale sensitivity.
# Each plot is displayed and saved to figs/ as a separate PDF.

source(here::here("src", "cdr_scale_sensitivity_visualisation.R"))

if (!exists("scale_delay_results")) {
  stop("scale_delay_results not found. Run cdr_scale_sensitivity_delay_analysis first.")
}

delay_keys <- grep("^delay_", names(scale_delay_results), value = TRUE)

for (dk in delay_keys) {
  delay_val <- as.integer(sub("^delay_", "", dk))
  cat(sprintf("\nPlotting outcome heatmap for delay = %d years...\n", delay_val))

  p <- create_cdr_scale_outcome_plot(
    sensitivity_results = scale_delay_results[[dk]],
    save_plot           = TRUE,
    filename            = sprintf("cdr_scale_outcome_delay_%d", delay_val),
    verbose             = FALSE
  )

  print(p)
}
```

```{r cdr_scale_sensitivity_visualisation, eval=FALSE}
# ==============================================================================
# CDR Scale Sensitivity Visualisation
# ==============================================================================
# Creates heatmap dashboards showing how CDR carrying capacity (K) and growth
# rate (r) affect climate outcomes across all SSP scenarios.
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "cdr_scale_sensitivity_visualisation.R"))

if (!exists("cdr_scale_sensitivity_results")) {
  stop("CDR scale sensitivity results not found. ",
       "Please run the cdr_scale_sensitivity_analysis chunk first.")
}

# Combined temperature heatmap (peak temperature and years above 1.5°C)
cdr_scale_dashboards_combined <- create_cdr_scale_temperature_dashboards_combined(
  sensitivity_results   = cdr_scale_sensitivity_results,
  add_contours          = TRUE,
  add_threshold_line    = FALSE, # NEEDS FIXING. LINE DIVIDES NO OVERSHOOT FROM RECOVERABLE
  threshold_value_temp  = 1.5,
  threshold_value_years = 0.5,
  show_infeasible       = FALSE,
  width                 = 297,
  height                = 140,
  save_plot             = TRUE,
  verbose               = TRUE
)

# Outcome plot distinguishing recoverable and unrecoverable parameter space
cdr_outcome_plot <- create_cdr_scale_outcome_plot(
  sensitivity_results = cdr_scale_sensitivity_results,
  threshold           = 1.505,
  alpha_recoverable   = 0.35,
  alpha_unrecoverable = 0.85,
  save_plot           = TRUE,
  width               = 297,
  height              = 100,
  verbose             = TRUE
)
```

```{r cdr_scale_data_results}
# ==============================================================================
# CDR Scale Sensitivity Data Extraction
# ==============================================================================
# Extracts three summary tables from the sensitivity results:
#   - feasible_min: lowest r and lowest K achieving recovery per SSP
#   - cdr_threshold: CDR units at the recoverable/unrecoverable boundary
#   - min_unrecoverable: peak temperature above which recovery is always impossible
# OUTPUTS: Saved automatically to output/ directory

source(here::here("src", "cdr_scale_data_extraction.R"))

if (!exists("cdr_scale_sensitivity_results")) {
  stop("CDR scale sensitivity results not found. ",
       "Please run the cdr_scale_sensitivity_analysis chunk first.")
}

# USER CHOICE: Thresholds defining a genuine overshoot and a feasible recovery
peak_temp_threshold  <- 1.51    # Minimum peak temperature to count as an overshoot
final_temp_threshold <- 1.501   # Maximum final temperature to count as recovered

cdr_scale_metrics <- extract_cdr_scale_metrics(
  sensitivity_results = cdr_scale_sensitivity_results,
  emissions_df        = emissions_df,
  peak_temp_min       = peak_temp_threshold,
  final_temp_max      = final_temp_threshold,
  g_initial           = g_initial,
  verbose             = TRUE
)
```

```{r free_time_scale_sensitivity_analysis, eval=FALSE}
# ==============================================================================
# CDR Scale Sensitivity — Free Terminal Time (Return Year Beyond 2100)
# ==============================================================================
# Free-terminal-time extension of the CDR scale sensitivity. For each K/r
# combination and scenario:
#   Stage 1: the standard fixed-time T = 2100 problem is solved. Cells that
#            meet the 1.5°C target by 2100 are recorded with T_star = 2100
#            (optimum_type = "feasible_by_2100", years_beyond_2100 = 0).
#   Stage 2: cells INFEASIBLE at 2100 get a free-terminal-time search over
#            (2100, 2500] for the earliest year the return to 1.5°C becomes
#            achievable (nested shooting: outer search on the terminal year
#            using the transversality Hamiltonian H(T), inner loop the usual
#            emissions-gap shooting). The headline output per cell is
#            years_beyond_2100 — how much longer that CDR scale needs.
#
# This chunk uses the EXTENDED-HORIZON data (emissions_df_2500/economic_df_2500,
# end_year = 2500) built by the data_preparation_2500_call chunk — genuine SSP
# projections beyond 2100, not flat-lined extrapolation — so the fixed-time
# chunks above (using emissions_df/economic_df, end_year = 2100) are unaffected.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row).
#   Run the data_preparation_2500_call chunk first.
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: Bimodal per cell — feasible-by-2100 cells cost one fixed solve;
#          infeasible cells run ~10-15 inner solves on horizons up to 2500.
#          ALWAYS run the runtime estimate below before a full grid: it times
#          a few representative cells and extrapolates.

source(here::here("src", "data_preparation.R"))
source(here::here("src", "optimal_control_capacity.R"))   # NOT optimal_control_core.R
source(here::here("src", "capacity_helpers.R"))
source(here::here("src", "optimal_control_free_time.R"))
source(here::here("src", "cdr_scale_sensitivity.R"))       # build_cdr_scale_grid()
source(here::here("src", "cdr_scale_free_time_sensitivity.R"))

if (!exists("emissions_df_2500") || !exists("economic_df_2500")) {
  stop("emissions_df_2500/economic_df_2500 not found. ",
       "Please run the data_preparation_2500_call chunk first.")
}

# USER CHOICE: CDR K range (carrying capacity, GtCO2/year)
K_min <- 25     # Minimum CDR carrying capacity
K_max <- 200    # Maximum CDR carrying capacity
n_K   <- 21     # Number of K values (use 11 for testing, 51 for publication)

# USER CHOICE: CDR r range (growth rate)
r_min       <- 0.02    # Minimum CDR growth rate
r_max       <- 0.20    # Maximum CDR growth rate
n_r         <- 21      # Number of r values (use 11 for testing, 51 for publication)
r_log_scale <- FALSE   # Set TRUE for log-spaced r values

# USER CHOICE: Fixed CDR starting conditions
g_initial <- 2      # Starting CDR deployment level (GtCO2/year)
t_start   <- 2025   # Year CDR deployment begins

# USER CHOICE: Scenarios to analyse (one SSP for testing, five for publication)
scenarios_to_compare <- c("SSP2-Baseline")

free_time_grid <- build_cdr_scale_grid(
  K_min       = K_min,
  K_max       = K_max,
  n_K         = n_K,
  r_min       = r_min,
  r_max       = r_max,
  n_r         = n_r,
  r_log_scale = r_log_scale
)

# ------------------------------------------------------------------------------
# Runtime estimate from a small timed subset — check this BEFORE the full run
# ------------------------------------------------------------------------------
free_time_runtime <- estimate_free_time_sweep_runtime(
  parameter_df = parameter_df[1, ],
  emissions_df = emissions_df_2500,
  economic_df  = economic_df_2500,
  scenarios    = scenarios_to_compare,
  cdr_grid     = free_time_grid,
  n_sample     = 6,
  g_initial    = g_initial,
  t_start      = t_start
)

# ------------------------------------------------------------------------------
# Full sweep (mitigation zeroed so CDR is the sole control, matching the
# fixed-time CDR scale sensitivity)
# ------------------------------------------------------------------------------
# checkpoint_path/progress_log are named from scenarios_to_compare + grid
# resolution so a different USER CHOICE configuration above never resumes
# from (and silently mixes with) an incompatible prior run's checkpoint.
# While this runs, tail progress from a separate session with (PowerShell):
#   Get-Content <path shown below> -Wait -Tail 20
run_tag <- paste0(paste(gsub("-Baseline", "", scenarios_to_compare), collapse = "_"),
                  "_", n_K, "x", n_r)

free_time_scale_results <- run_free_time_scale_sensitivity(
  parameter_df                  = parameter_df[1, ],
  emissions_df                  = emissions_df_2500,
  economic_df                   = economic_df_2500,
  scenarios                     = scenarios_to_compare,
  cdr_grid                      = free_time_grid,
  g_initial                     = g_initial,
  t_start                       = t_start,
  t_max                         = 2500,
  use_mitigation_capacity_limit = TRUE,
  mitigation_capacity_function  = make_zero_capacity(),
  use_parallel                  = TRUE,
  checkpoint_path               = here::here("output", paste0("checkpoint_free_time_2500_", run_tag, ".rds")),
  progress_log                  = here::here("output", paste0("progress_free_time_2500_", run_tag, ".log")),
  save_results                  = TRUE,
  verbose                       = TRUE
)

# Quick look: how many extra years past 2100 does each infeasible cell need?
with(free_time_scale_results$combined_results, {
  cat("Feasible by 2100:      ", sum(optimum_type == "feasible_by_2100", na.rm = TRUE), "\n")
  cat("Return in 2101-2500:   ", sum(feasible & T_star > 2100, na.rm = TRUE), "\n")
  cat("No return by 2500:     ", sum(!feasible), "\n")
})
```

```{r free_time_scale_sensitivity_visualisation, eval=FALSE}
# ==============================================================================
# CDR Scale Free-Terminal-Time Visualisation — Return Year Heatmap
# ==============================================================================
# Heatmap of the return year T_star across the K/r grid, one panel per SSP
# scenario present in the results. Colour scale is fixed to [2100, 2300] (the
# free-time search horizon) rather than data-driven, so shading is directly
# comparable across scenarios and figures. Cells with no return achieved
# within the horizon (feasible = FALSE) are shown grey with a red "x".
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "cdr_scale_sensitivity_visualisation.R"))  # shared theme/helpers
source(here::here("src", "cdr_scale_free_time_visualisation.R"))

if (!exists("free_time_scale_results")) {
  stop("free_time_scale_results not found. ",
       "Please run the free_time_scale_sensitivity_analysis chunk first.")
}

free_time_year_heatmap <- create_free_time_year_heatmap(
  sensitivity_results = free_time_scale_results,
  year_limits         = c(2100, 2300),
  show_infeasible     = TRUE,
  save_plot           = TRUE,
  width               = 297,
  height              = 100,
  verbose             = TRUE
)

print(free_time_year_heatmap)
```


```{r scenario_comparison_analysis_capacity, eval=FALSE}
# ==============================================================================
# Scenario Comparison Analysis with Capacity Constraints
# ==============================================================================
# Compares optimal control solutions across SSP scenarios with realistic
# deployment constraints enforced on mitigation and CDR via logistic curves.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: ~5-15 minutes for 5 scenarios with parallel processing

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))

# USER CHOICE: SSP scenarios to compare
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline",
                          "SSP4-Baseline", "SSP5-Baseline")

# USER CHOICE: Delay settings (typically keep at 0 for baseline comparison)
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay        <- 0  # Years before CDR deployment begins

# USER CHOICE: Mitigation capacity constraint
# Set use_mitigation_capacity_limit = FALSE to allow unconstrained mitigation.
# make_zero_capacity() eliminates mitigation so CDR is the sole control.
use_mitigation_capacity_limit <- TRUE
mitigation_capacity_function  <- make_zero_capacity()

# USER CHOICE: CDR capacity constraint
# Logistic S-curve starting from current CDR levels and growing to K (GtCO2/year).
# Increase r for faster ramp-up; increase K for a higher ceiling.
use_cdr_capacity_limit <- TRUE
cdr_capacity_function  <- make_logistic_from_zero(
  g_initial = 2,      # Current CDR deployment level (GtCO2/year)
  K         = 100,    # Maximum CDR capacity (GtCO2/year)
  r         = 0.07,   # Growth rate (0.05 = slow, 0.10 = fast)
  t_start   = 2025
)

# Run scenario comparison
scenario_comparison_results <- run_scenario_comparison(
  parameter_df                  = parameter_df[1, ],
  emissions_df                  = emissions_df,
  economic_df                   = economic_df,
  scenarios                     = scenarios_to_compare,
  mitigation_delay_years        = mitigation_delay,
  cdr_delay_years               = cdr_delay,
  use_mitigation_capacity_limit = use_mitigation_capacity_limit,
  mitigation_capacity_function  = mitigation_capacity_function,
  use_cdr_capacity_limit        = use_cdr_capacity_limit,
  cdr_capacity_function         = cdr_capacity_function,
  use_parallel                  = TRUE,
  save_results                  = TRUE,
  verbose                       = FALSE
)

if (!is.null(scenario_comparison_results$comparison_summary)) {
  print(scenario_comparison_results$comparison_summary)
}
```

```{r multi_scenario_comparison_analysis_capacity, eval=FALSE}
# ==============================================================================
# Multi-Growth-Rate Scenario Comparison with Capacity Constraints
# ==============================================================================
# Sweeps over a set of named CDR growth rates (r), running the full SSP scenario
# comparison once per rate. Results are saved as per-rate and combined RDS files.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: ~5-15 minutes per growth rate (e.g. 3 rates × 5 SSPs ≈ 15-45 min)

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))

# USER CHOICE: SSP scenarios to compare
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline",
                          "SSP4-Baseline", "SSP5-Baseline")

# USER CHOICE: Delay settings (typically keep at 0 for baseline comparison)
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay        <- 0  # Years before CDR deployment begins

# USER CHOICE: CDR growth rates to compare
# Each entry is a named r value. The name becomes the key in results and filenames.
# Common values: slow ≈ 0.05, moderate ≈ 0.07, fast ≈ 0.11
cdr_r_values <- c(
  slow     = 0.05,
  moderate = 0.07,
  fast     = 0.11
)

# USER CHOICE: Shared CDR logistic parameters (fixed across all growth rates)
cdr_g_initial <- 2     # Starting CDR deployment level (GtCO2/year)
cdr_K         <- 100   # Maximum CDR carrying capacity (GtCO2/year)
cdr_t_start   <- 2025  # Year CDR deployment begins

# USER CHOICE: Mitigation capacity constraint
# make_zero_capacity() eliminates mitigation so CDR is the sole control.
use_mitigation_capacity_limit <- TRUE
mitigation_capacity_function  <- make_zero_capacity()

# Run comparison across all growth rates
capacity_growth_results <- run_capacity_growth_comparison(
  parameter_df                  = parameter_df[1, ],
  emissions_df                  = emissions_df,
  economic_df                   = economic_df,
  scenarios                     = scenarios_to_compare,
  r_values                      = cdr_r_values,
  g_initial                     = cdr_g_initial,
  K                             = cdr_K,
  t_start                       = cdr_t_start,
  mitigation_delay_years        = mitigation_delay,
  cdr_delay_years               = cdr_delay,
  use_mitigation_capacity_limit = use_mitigation_capacity_limit,
  mitigation_capacity_function  = mitigation_capacity_function,
  use_parallel                  = TRUE,
  save_results                  = TRUE,
  verbose                       = TRUE
)

# Display summary across all growth rates
for (rate in names(capacity_growth_results)) {
  cat("\n--- Growth rate:", rate, "---\n")
  summary_df <- capacity_growth_results[[rate]]$comparison_summary
  if (!is.null(summary_df)) {
    print(summary_df[, c("scenario", "peak_temperature", "total_cost",
                         "total_cdr_units", "years_above_1p5")])
  }
}
```

```{r lowest_comparison_analysis, eval=FALSE}
# ==============================================================================
# Lowest Comparison Analysis
# ==============================================================================
# Runs optimal control for SSP-specific (r, K) parameter pairs. Unlike the
# multi-growth-rate chunk above, each SSP receives its own set of exact (r, K)
# combinations rather than a shared sweep — useful for targeted analysis of
# minimum feasible CDR parameters identified from the sensitivity grid.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ as a single combined RDS
# RUNTIME: ~5-15 minutes per (SSP, pair) with parallel processing

source(here::here("src", "lowest_comparison.R"))
source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))

# USER CHOICE: SSP-specific (r, K) parameter pairs
# Each SSP name must match a scenario in emissions_df exactly.
# Each SSP receives a list of c(r = ..., K = ...) vectors, one per run.
# Pairs are labelled automatically as pair_1, pair_2, ...
# K must be greater than cdr_g_initial for every pair.
ssp_params <- list(
  "SSP1-Baseline" = list(
    c(r = 0.045, K = 95),
    c(r = 0.07,  K = 25)
  ),
  "SSP2-Baseline" = list(
    c(r = 0.06,  K = 161.5),
    c(r = 0.11,  K = 42.5)
  ),
  "SSP3-Baseline" = list(
    c(r = 0.063, K = 144),
    c(r = 0.128, K = 42.5)
  ),
  "SSP4-Baseline" = list(
    c(r = 0.052, K = 182.5),
    c(r = 0.189, K = 25)
  ),
  "SSP5-Baseline" = list(
    c(r = 0.074, K = 186),
    c(r = 0.157, K = 63.5)
  )
)

# USER CHOICE: Shared CDR starting conditions (fixed across all SSPs and pairs)
cdr_g_initial <- 2     # Starting CDR deployment level (GtCO2/year)
cdr_t_start   <- 2025  # Year CDR deployment begins

# USER CHOICE: Delay settings
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay        <- 0  # Years before CDR deployment begins

# USER CHOICE: Mitigation capacity constraint
# make_zero_capacity() eliminates mitigation so CDR is the sole control.
use_mitigation_capacity_limit <- TRUE
mitigation_capacity_function  <- make_zero_capacity()

# Run analysis across all SSPs and pairs
lowest_results <- run_lowest_comparison(
  parameter_df                  = parameter_df[1, ],
  emissions_df                  = emissions_df,
  economic_df                   = economic_df,
  ssp_params                    = ssp_params,
  g_initial                     = cdr_g_initial,
  t_start                       = cdr_t_start,
  mitigation_delay_years        = mitigation_delay,
  cdr_delay_years               = cdr_delay,
  use_mitigation_capacity_limit = use_mitigation_capacity_limit,
  mitigation_capacity_function  = mitigation_capacity_function,
  use_parallel                  = TRUE,
  save_results                  = TRUE,
  verbose                       = TRUE
)

# Display summary across all SSPs and pairs
for (ssp in names(lowest_results)) {
  cat("\n---", ssp, "---\n")
  for (pair in names(lowest_results[[ssp]])) {
    run <- lowest_results[[ssp]][[pair]]
    summary_df <- run$comparison_summary
    if (!is.null(summary_df)) {
      cat("  ", pair, "| peak temp:", round(summary_df$peak_temperature, 2),
          "°C | years >1.5°C:", summary_df$years_above_1p5,
          "| total cost:", round(summary_df$total_cost, 1), "\n")
    }
  }
}
```

```{r lowest_comparison_visualisation, eval=FALSE}
# ==============================================================================
# Lowest Comparison Visualisation
# ==============================================================================
# Creates a 2-row × N-column dashboard comparing results across all SSPs and
# parameter pairs: CDR deployment (top row) and temperature trajectories
# (bottom row). pair_1 and pair_2 are shown as different linetypes.
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "lowest_analysis_dashboard.R"))

if (!exists("lowest_results")) {
  stop("lowest_results not found. Please run the lowest_comparison_analysis chunk first.")
}

# USER CHOICE: Dashboard display options
# show_capacity_limit = TRUE overlays the CDR capacity curve on the CDR panel.
# Pass the same ssp_params, cdr_g_initial, and cdr_t_start used during analysis
# so capacity curves can be reconstructed (even when reloading from an RDS file).
lowest_dashboard <- create_lowest_comparison_dashboard(
  lowest_results      = lowest_results,
  show_capacity_limit = TRUE,
  ssp_params          = ssp_params,
  g_initial           = cdr_g_initial,
  t_start             = cdr_t_start,
  save_plot           = TRUE,
  width               = 297,   # A4 landscape width (mm)
  height              = 210,
  verbose             = TRUE
)

print(lowest_dashboard)
```

```{r extract_capacity_summary, eval=FALSE}
# ==============================================================================
# Extract Capacity Growth Summary
# ==============================================================================
# Pulls peak temperature and years above 1.5°C for every SSP × growth rate
# combination into a tidy data frame and saves it as a timestamped CSV.
#
# REQUIRES: capacity_growth_results (from multi_scenario_comparison_analysis_capacity)

source(here::here("src", "capacity_helpers.R"))

if (!exists("capacity_growth_results")) {
  stop("capacity_growth_results not found. ",
       "Please run the multi_scenario_comparison_analysis_capacity chunk first.")
}

capacity_summary_df <- extract_capacity_summary(
  capacity_growth_results = capacity_growth_results,
  verbose                 = TRUE
)

print(capacity_summary_df)
```

```{r scenario_comparison_visualisation_capacity, eval=FALSE}
# ==============================================================================
# Capacity Scenario Comparison Visualisation
# ==============================================================================
# Creates a 2-row × 5-column SSP-faceted figure comparing temperature
# trajectories and CDR deployment across growth rates.
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "scenario_comparison_capacity_visualisation.R"))

if (!exists("capacity_growth_results")) {
  stop("Capacity growth results not found. ",
       "Please run the multi_scenario_comparison_analysis_capacity chunk first.")
}

# USER CHOICE: Dashboard display options
capacity_dashboard <- create_capacity_scenario_comparison_dashboard(
  capacity_results    = capacity_growth_results,
  show_capacity_limit = TRUE,
  save_plot           = TRUE,
  width               = 297,   # A4 landscape width (mm)
  height              = 210,
  verbose             = TRUE
)

print(capacity_dashboard)
```


```{r cdr_delay_growth_analysis, eval=FALSE}
# ==============================================================================
# CDR Delay x Growth Rate Analysis
# ==============================================================================
# Sweeps CDR deployment delay (0 to max_delay_years) across a set of named CDR
# growth rates (slow, moderate, fast) with a fixed carrying capacity K.
# Mitigation is eliminated so CDR is the sole control throughout.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Intermediate per-rate RDS files + combined RDS saved to output/
# RUNTIME: Scales with length(r_values) x (max_delay_years / delay_step_size)
#          Approx 0.27 min per scenario comparison (5 SSPs in parallel):
#          step_size=1  (~213 comparisons) ≈ 57 min
#          step_size=5  (~45 comparisons)  ≈ 12 min
#          step_size=10 (~24 comparisons)  ≈ 6.5 min

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))
source(here::here("src", "cdr_delay_growth.R"))

# USER CHOICE: CDR growth rates to compare (named vector)
# Common values: slow ≈ 0.05, moderate ≈ 0.07, fast ≈ 0.11
cdr_r_values <- c(
  slow     = 0.05,
  moderate = 0.07,
  fast     = 0.11
)

# USER CHOICE: Fixed CDR logistic parameters (shared across all growth rates)
cdr_K         <- 100   # Maximum CDR carrying capacity (GtCO2/year)
cdr_g_initial <- 2     # Starting CDR deployment level (GtCO2/year)
cdr_t_start   <- 2025  # Year CDR deployment begins

# USER CHOICE: Delay range and step size
max_delay_years <- 70   # Maximum CDR delay to test (years)
delay_step_size <- 10   # Step size — set to 1 for publication-quality results,
                        # larger for fast exploratory runs (e.g. 5 or 10)

# USER CHOICE: Scenarios to analyse
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline",
                          "SSP4-Baseline", "SSP5-Baseline")

cdr_delay_growth_results <- run_cdr_delay_growth_analysis(
  parameter_df                  = parameter_df[1, ],
  emissions_df                  = emissions_df,
  economic_df                   = economic_df,
  scenarios                     = scenarios_to_compare,
  r_values                      = cdr_r_values,
  K                             = cdr_K,
  g_initial                     = cdr_g_initial,
  t_start                       = cdr_t_start,
  max_delay_years               = max_delay_years,
  delay_step_size               = delay_step_size,
  use_mitigation_capacity_limit = TRUE,
  mitigation_capacity_function  = make_zero_capacity(),
  use_parallel                  = TRUE,
  save_results                  = TRUE,
  verbose                       = TRUE
)
```

```{r cdr_delay_growth_outcome_plots, eval=FALSE}
# ==============================================================================
# CDR Delay x Growth Rate Outcome Plots
# ==============================================================================
# Three plots using the same tidy summary data:
#   1. Line plot — peak temperature vs CDR delay (one line per growth rate,
#      5 SSP column facets)
#   2. Line plot — years above 1.5C vs CDR delay (same structure)
#   3. Threshold dot plot — last feasible CDR delay before outcome becomes
#      unrecoverable, by SSP and growth rate
#
# Each saved as a separate PDF and PNG to figs/

source(here::here("src", "cdr_delay_growth_visualisation.R"))

if (!exists("cdr_delay_growth_results")) {
  stop("Results not found. Please run the cdr_delay_growth_analysis chunk first.")
}

# USER CHOICE: Thresholds (match cdr_scale_data_results for consistency)
peak_temp_threshold  <- 1.51
final_temp_threshold <- 1.501

# Extract tidy summary (shared by all three plots)
if (!exists("delay_summary_df")) {
  delay_summary_df <- extract_delay_summary(
    delay_growth_results = cdr_delay_growth_results,
    peak_temp_threshold  = peak_temp_threshold,
    final_temp_threshold = final_temp_threshold
  )
}

# 1. Peak temperature vs delay
plot_peak_temp <- create_delay_line_plot(
  summary_df = delay_summary_df,
  variable   = "peak_temperature",
  save_plot  = TRUE,
  width      = 297,
  height     = 100
)
print(plot_peak_temp)

# 2. Years above 1.5C vs delay
plot_years <- create_delay_line_plot(
  summary_df = delay_summary_df,
  variable   = "years_above_1p5",
  save_plot  = TRUE,
  width      = 297,
  height     = 100
)
print(plot_years)

# 3. Last feasible delay threshold
plot_threshold <- create_delay_threshold_plot(
  summary_df = delay_summary_df,
  save_plot  = TRUE,
  width      = 180,
  height     = 100
)
print(plot_threshold)
```

```{r cdr_delay_growth_heatmaps, eval=FALSE}
# ==============================================================================
# CDR Delay x Growth Rate Heatmaps
# ==============================================================================
# Creates three heatmaps sharing the same axes:
#   x = CDR delay (years), y = growth rate (slow/moderate/fast),
#   facets = SSP scenario (5 columns)
#
#   1. Outcome heatmap: categorical fill (no overshoot / recoverable /
#      unrecoverable overshoot)
#   2. Peak temperature heatmap: continuous fill
#   3. Years above 1.5C heatmap: continuous fill
#
# Each is saved as a separate PDF and PNG to figs/

source(here::here("src", "cdr_delay_growth_visualisation.R"))

if (!exists("cdr_delay_growth_results")) {
  stop("Results not found. Please run the cdr_delay_growth_analysis chunk first.")
}

# USER CHOICE: Thresholds defining overshoot and recovery (match cdr_scale_data_results)
peak_temp_threshold  <- 1.51    # Minimum peak temperature to count as overshoot
final_temp_threshold <- 1.501   # Maximum final temperature to count as recovered

# Extract tidy summary data frame (computed once, used by all three heatmaps)
delay_summary_df <- extract_delay_summary(
  delay_growth_results = cdr_delay_growth_results,
  peak_temp_threshold  = peak_temp_threshold,
  final_temp_threshold = final_temp_threshold
)

# 1. Outcome heatmap (categorical)
heatmap_outcome <- create_delay_heatmap(
  summary_df = delay_summary_df,
  variable   = "outcome",
  save_plot  = TRUE,
  width      = 297,
  height     = 100
)
print(heatmap_outcome)

# 2. Peak temperature heatmap (continuous)
heatmap_peak_temp <- create_delay_heatmap(
  summary_df = delay_summary_df,
  variable   = "peak_temperature",
  save_plot  = TRUE,
  width      = 297,
  height     = 100
)
print(heatmap_peak_temp)

# 3. Years above 1.5C heatmap (continuous)
heatmap_years <- create_delay_heatmap(
  summary_df = delay_summary_df,
  variable   = "years_above_1p5",
  save_plot  = TRUE,
  width      = 297,
  height     = 100
)
print(heatmap_years)
```

```{r cdr_delay_growth_visualisation, eval=FALSE}
# ==============================================================================
# CDR Delay x Growth Rate Visualisation
# ==============================================================================
# Creates a line plot dashboard showing CDR deployment and temperature
# trajectories across SSP scenarios. Lines within each panel are coloured by
# CDR delay (plasma colour scale: dark = short delay, bright = long delay).
# Growth rates are separated into row groups; SSPs are columns.
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "cdr_delay_growth_visualisation.R"))

if (!exists("cdr_delay_growth_results")) {
  stop("Results not found. Please run the cdr_delay_growth_analysis chunk first.")
}

# USER CHOICE: Which growth rates to include
# Set to NULL to include all rates, or a subset e.g. c("slow", "moderate")
rates_to_plot <- NULL

# USER CHOICE: Visual options
show_cap_limit <- TRUE   # Overlay CDR capacity ceiling on CDR panels
alpha_lines    <- 0.7    # Line transparency — lower if many lines overlap

cdr_delay_growth_dashboard <- create_cdr_delay_growth_dashboard(
  delay_growth_results = cdr_delay_growth_results,
  rates                = rates_to_plot,
  show_capacity_limit  = show_cap_limit,
  line_alpha           = alpha_lines,
  save_plot            = TRUE,
  width                = 297,
  height               = NULL,   # Auto: 140mm per growth rate included
  verbose              = TRUE
)

print(cdr_delay_growth_dashboard)
```

```{r delayed_deployment_capacity_analysis, eval=FALSE}
# ==============================================================================
# Delayed Deployment Analysis with CDR Capacity Constraint
# ==============================================================================
# Combines the delayed_deployment_analysis grid (mitigation_delay x cdr_delay)
# with the CDR logistic capacity cap from the capacity-constrained scenario
# comparison. Mitigation remains a free, unconstrained control (only delayed);
# CDR is both delayed AND capped by a logistic growth curve, so its ramp-up
# speed is limited by the chosen capacity level (slow/moderate/fast).
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Per-capacity-level RDS + combined RDS/CSV saved to output/
# RUNTIME: No bespoke fast solver exists for delay x capacity combinations, so
#          this uses the general capacity-aware solver and is much slower per
#          combination than delayed_deployment_analysis. Measured: ~0.175 min
#          per scenario comparison (5 SSPs in parallel):
#          step_size=20, max_delay=20 (2x2x3=12 comparisons)   => 2.1 min (measured)
#          step_size=10, max_delay=70 (8x8x3=192 comparisons)  => 33.3 min (measured)
#          step_size=5,  max_delay=70 (15x15x3=675 comparisons) => ~2 hrs (est.)
#          step_size=1,  max_delay=70 (71x71x3=15123 comparisons) => ~1.8 days, do not use
#          Use a larger step size for exploratory runs; only drop to smaller
#          step sizes for final publication figures with time to spare.

source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))
source(here::here("src", "delayed_deployment_capacity.R"))

# USER CHOICE: Delay range and granularity (applies to BOTH mitigation and CDR delay)
max_delay_years <- 70        # Maximum delay to test (years)
delay_step_size <- 10        # Step size — larger for testing, 1 for publication (see runtime note above)

# USER CHOICE: Scenarios to analyse
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline",
                          "SSP4-Baseline", "SSP5-Baseline")

# USER CHOICE: Named CDR growth rates defining the capacity levels to compare
# Common values: slow ≈ 0.05, moderate ≈ 0.07, fast ≈ 0.11
cdr_r_values <- c(
  slow     = 0.05,
  moderate = 0.07,
  fast     = 0.11
)

# USER CHOICE: Shared CDR logistic parameters (fixed across all capacity levels)
cdr_g_initial <- 2     # Starting CDR deployment level (GtCO2/year)
cdr_K         <- 100   # Maximum CDR carrying capacity (GtCO2/year)
cdr_t_start   <- 2025  # Year CDR deployment nominally begins

delayed_deployment_capacity_results <- run_delayed_deployment_capacity_analysis(
  parameter_df    = parameter_df[1, ],
  emissions_df    = emissions_df,
  economic_df     = economic_df,
  scenarios       = scenarios_to_compare,
  r_values        = cdr_r_values,
  K               = cdr_K,
  g_initial       = cdr_g_initial,
  t_start         = cdr_t_start,
  max_delay_years = max_delay_years,
  delay_step_size = delay_step_size,
  use_parallel    = TRUE,
  save_results    = TRUE,
  verbose         = TRUE
)

# Display summary
cat("\n=== DELAYED DEPLOYMENT + CAPACITY ANALYSIS COMPLETE ===\n")
cat("Delay range tested: 0 -", max_delay_years, "years (step:", delay_step_size, ")\n")
cat("Capacity levels:", paste(names(cdr_r_values), collapse = ", "), "\n")
cat("Total rows:", nrow(delayed_deployment_capacity_results$combined_results), "\n")
```

```{r delayed_deployment_capacity_visualisation, eval=FALSE}
# ==============================================================================
# Delayed Deployment with CDR Capacity — Visualisation
# ==============================================================================
# Reuses the same heatmap dashboard function as delayed_deployment_visualisation,
# producing one dashboard per CDR capacity level (slow/moderate/fast). Each
# dashboard is placed on its own page within a single multi-page PDF.
# OUTPUTS: Saved automatically to figs/ directory

source(here::here("src", "delayed_deployment_visualisation.R"))

if (!exists("delayed_deployment_capacity_results")) {
  stop("delayed_deployment_capacity_results not found. Please run the delayed_deployment_capacity_analysis chunk first.")
}

# Create abatement cost field (mitig cost + remov_cost)
delayed_deployment_capacity_results$combined_results <- delayed_deployment_capacity_results$combined_results %>%
  mutate(abatement_cost = mitig_cost + remov_cost)

capacity_levels <- names(delayed_deployment_capacity_results$results_by_capacity)

# USER CHOICE: Variables to plot (same options as delayed_deployment_visualisation)
dashboard_variables <- c("peak_temperature", "years_above_1p5")

# USER CHOICE: Output PDF dimensions (A4 landscape, mm)
pdf_width  <- 297
pdf_height <- 210

# Compute colour scale limits ONCE across ALL capacity levels (not per-page)
# so slow/moderate/fast use the exact same scale and are directly comparable
# at a glance. shared_scale stays FALSE here too -- peak_temperature (°C) and
# years_above_1p5 (years) still get their own independent scale from each
# other; what's now shared is each variable's own scale ACROSS the 3 pages.
global_variable_limits <- calculate_variable_limits(
  data = delayed_deployment_capacity_results$combined_results,
  variables = dashboard_variables,
  shared_scale = FALSE,
  use_scale_limits = TRUE,
  scale_limit_percentile = 70,
  verbose = FALSE
)

pdf_filename <- paste0("delayed_deployment_capacity_dashboard_",
                       format(Sys.time(), "%Y%m%d_%H%M%S"), ".pdf")
pdf_path <- here::here("figs", pdf_filename)

cairo_pdf(pdf_path, width = pdf_width / 25.4, height = pdf_height / 25.4, onefile = TRUE)

for (level in capacity_levels) {

  level_data <- delayed_deployment_capacity_results$combined_results %>%
    filter(capacity_level == level)

  dashboard <- create_delayed_deployment_dashboard(
    deployment_results = level_data,
    variables = dashboard_variables,
    add_contours = TRUE,
    contour_alpha = 0.6,
    add_arrows = FALSE,
    show_infeasible = FALSE,
    color_palettes = NULL,
    # USER CHOICE: shared_scale should only be TRUE when all requested variables
    # share the same units (e.g. the three cost variables in $ trillion).
    # peak_temperature (°C) and years_above_1p5 (years) do NOT share units, so
    # this must stay FALSE -- each row gets its own accurately-scoped legend.
    shared_scale = FALSE,
    variable_limits_override = global_variable_limits,
    save_plot = FALSE,
    verbose = FALSE
  )

  # Label each page with which capacity level it shows (create_delayed_deployment_
  # dashboard()'s own title/subtitle are disabled project-wide, so add it directly)
  dashboard <- dashboard +
    patchwork::plot_annotation(
      title = paste0("CDR Capacity: ", tools::toTitleCase(level)),
      theme = theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
    )

  print(dashboard)
}

dev.off()

cat("Saved multi-page dashboard to:", pdf_path, "\n")
```

```{r test}

library(here); library(dplyr); library(ggplot2); library(patchwork); library(scales)

source(here::here("src", "cdr_scale_sensitivity_visualisation.R"))

cdr_scale_sensitivity_results <- readRDS(
  here::here("output", "cdr_scale_free_time_sensitivity_20260825_195408.rds")
)

cdr_scale_dashboards_combined <- create_cdr_scale_temperature_dashboards_combined(
  sensitivity_results   = cdr_scale_sensitivity_results,
  add_contours          = FALSE,
  add_threshold_line    = FALSE,
  threshold_value_temp  = 1.5,
  threshold_value_years = 0.5,
  show_infeasible       = TRUE,
  width                 = 297,
  height                = 140,
  save_plot             = TRUE,
  filename              = "cdr_scale_temperature_combined_freetime_2500_SSP1_21x21",
  verbose               = TRUE
)

print(cdr_scale_dashboards_combined)
```

```{r test2}

library(here); library(dplyr); library(ggplot2); library(patchwork); library(scales)

source(here::here("src", "cdr_scale_sensitivity_visualisation.R"))
source(here::here("src", "cdr_scale_free_time_visualisation.R"))

free_time_scale_results <- readRDS(
  here::here("output", "cdr_scale_free_time_sensitivity_20260825_195408.rds")
)

free_time_year_heatmap <- create_free_time_year_heatmap(
  sensitivity_results = free_time_scale_results,
  year_limits         = c(2101, 2500),
  show_infeasible     = TRUE,
  save_plot           = TRUE,
  width               = 297,
  height              = 100,
  verbose             = TRUE
)

print(free_time_year_heatmap)


```

# ==============================================================================
# Workflow Complete
# ==============================================================================
# 
# If all chunks have run successfully, you should now have:
# ✓ Processed emissions and economic data for all SSP scenarios
# ✓ Generated parameter sets (either fixed or LHS)
# ✓ Optimal control solutions for selected scenarios
# ✓ Publication-ready visualisations in figs/ directory
# ✓ Results files saved in output/ directory
#
# NEXT STEPS:
# - Load saved RDS files for detailed analysis
# - Customise visualisations by modifying plotting parameters
# - Run different parameter sets or scenarios
# - See USER_GUIDE.md for advanced usage and troubleshooting
#
# GETTING HELP:
# - Email: nina.rynne@griffithuni.edu.au
# - Repository: https://github.com/nina-rynne/MACROM
# - Issues: Submit via GitHub issue tracker
#
# Thank you for using MACROM!
# ==============================================================================