---
title: "MACROM: Optimal Control Workflow for Climate Temperature Overshoot"
author: "Nina Rynne"
institution: "Griffith University"
output: html_document
date: "January 2026"
---

# ==============================================================================
# MACROM: An Optimal Control Model for Balancing Climate Change 
# Abatement and Damage Trade-offs
# 
# Authors: Nina Rynne, Michael Bode, Melanie Roberts, Ryan Heneghan
# Institution: Griffith University
# 
# Copyright (c) 2025 Nina Rynne
# Licensed under CC-BY-4.0 - see LICENSE file for details
# 
# Citation: If you use this code, please cite:
#   Rynne, N., Bode, M., Roberts, M., & Heneghan, R. (2025). 
#   MACROM: An Optimal Control Model for Balancing Climate Change 
#   Abatement and Damage Trade-offs.
#   https://doi.org/10.5281/zenodo.18463951
# 
# Version: 1.0.0
# Last updated: January 2026
# 
# Repository: https://github.com/nina-rynne/MACROM
# Contact: nina.rynne@griffithuni.edu.au
# ==============================================================================

# ABOUT MACROM:
# This workflow runs the complete MACROM optimal control model. MACROM determines
# optimal climate mitigation and CDR strategies by balancing abatement costs 
# against temperature-related damages under different SSP emission scenarios.
#
# This is the primary user interface - public users interact only with this file.
# For detailed instructions, prerequisites, and troubleshooting, see USER_GUIDE.md

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Create output directories if they don't exist
dir.create(here::here("output"), showWarnings = FALSE)
dir.create(here::here("figs"), showWarnings = FALSE)

```

```{r libraries}

# ==============================================================================
# Required Libraries
# ==============================================================================
# Install any missing packages using: install.packages("package_name")

# ============================================
# Data Manipulation and Transformation
# ============================================
#library(dplyr)        # Data manipulation
#library(tidyr)        # Data reshaping
library(purrr)        # Functional programming tools
library(tidyverse)    # Data science toolkit

# ============================================
# File Handling and I/O
# ============================================
library(readr)        # Reading data files
library(here)         # File path management
library(yaml)         # YAML file handling
library(Cairo)        # PDF export

# ============================================
# Parallel Processing
# ============================================
library(parallel)     # Parallel processing
library(foreach)      # Parallel iteration
library(doParallel)   # Parallel backend

# ============================================
# Data Analysis and Statistics
# ============================================
library(lhs)          # Latin hypercube sampling
library(iterators)    # Iterator support

# ============================================
# Visualisation
# ============================================
library(ggplot2)      # Data visualisation
library(cowplot)      # Plot arrangement
library(patchwork)    # Plot composition
library(viridisLite)  # Colourblind friendly palettes
library(viridis)      # Colourblind friendly palettes
library(colorspace)   # Advanced colour manipulation
library(scales)       # Scale functions for visualisation
library(metR)         # Contour line labelling
```

```{r data_preparation_call}
# ==============================================================================
# Data Preparation
# ==============================================================================
# Import and prepare emissions and economic data
# This section imports raw SSP scenario data and interpolates it to annual time steps

# Source data preparation functions
source(here::here("src", "data_preparation.R"))

# Import emissions and economic data
emissions_imported <- import_ssp_emissions("emissions.csv")
economic_imported <- import_ssp_economic("gwp.csv")

# Interpolate data to annual time step (2020-2100)
# USER CHOICE: Modify start_year, end_year, or dt if different time range/step is needed
emissions_df <- interpolate_ssp_emissions(
  emissions_df = emissions_imported,
  dt = 1,
  start_year = 2020,
  end_year = 2100
)

economic_df <- interpolate_ssp_economic(
  economic_df = economic_imported,
  dt = 1,
  start_year = 2020,
  end_year = 2100
)
```

```{r latin_hypercube_sampling_call, eval=FALSE}
# ==============================================================================
# PARAMETER APPROACH 1: Multiple Parameter Sets (Sensitivity Analysis)
# ==============================================================================
# ⚠️ USER CHOICE: Use EITHER this chunk OR the fixed_model_parameters_call chunk below
# Do NOT run both - they are mutually exclusive options
#
# This approach generates multiple parameter combinations for uncertainty quantification
# Parameter ranges defined in: parameter_details.yml
#
# WHEN TO USE:
# - Exploring parameter sensitivity
# - Uncertainty quantification
# - Generating confidence intervals
#
# NOTE: Significantly increases computation time
# Recommended: 100-500 samples for thorough analysis

source(here::here("src", "latin_hypercube_sampling.R"))
source(here::here("src", "model_parameters.R"))

# USER CHOICE: Change to desired number of parameter sets to generate
# More samples = better coverage but longer runtime
lhs_parameter_df <- generate_lhs_samples(n_samples = 100, seed = 12345)

# Add fixed model parameters to the LHS samples
parameter_df <- add_fixed_parameters(lhs_parameter_df)
```

```{r fixed_model_parameters_call}
# ==============================================================================
# PARAMETER APPROACH 2: Single Parameter Set (Recommended for most analyses)
# ==============================================================================
# ⚠️ USER CHOICE: Use EITHER this chunk OR the latin_hypercube_sampling_call chunk above
# Do NOT run both - they are mutually exclusive options
#
# This approach uses a single predefined parameter set
# Parameter values defined in: parameter_details_fixed.yml
#
# WHEN TO USE:
# - Comparing across SSP scenarios
# - Analysing delayed deployment effects
# - Generating primary results with best-estimate parameters
# - Quick test runs
#
# RECOMMENDED FOR: Most standard analyses and publication figures

source(here::here("src", "model_parameters.R"))

# Use predefined parameter set
parameter_df <- create_params_dataframe()
```

```{r scenario_comparison_analysis, eval=FALSE}
# ==============================================================================
# Scenario Comparison Analysis
# ==============================================================================
# Compare optimal control solutions across different SSP baseline scenarios
# 
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: ~5-15 minutes for 5 scenarios with parallel processing

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "scenario_comparison.R"))

# USER CHOICE: Select which SSP scenarios to compare
scenarios_to_compare <- c("SSP5-Baseline")

# USER CHOICE: Delay settings (typically keep at 0 for baseline comparison)
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay <- 0         # Years before CDR deployment begins

# Run scenario comparison
scenario_comparison_results <- run_scenario_comparison(
  parameter_df = parameter_df[1, ],      # Must be single row
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenarios = scenarios_to_compare,
  mitigation_delay_years = mitigation_delay,
  cdr_delay_years = cdr_delay,
  use_parallel = TRUE,                   # Set FALSE if parallel issues occur
  save_results = TRUE,                   # Automatically saves to output/
  verbose = FALSE                        # Set TRUE for detailed progress
)

# Display summary
if (!is.null(scenario_comparison_results$comparison_summary)) {
  print(scenario_comparison_results$comparison_summary)
}
```

```{r scenario_comparison_visualisation, eval=FALSE}
# ==============================================================================
# Scenario Comparison Visualisation
# ==============================================================================
# Create publication-quality visualisations comparing optimal control solutions
# OUTPUTS: Saved automatically to figs/ directory

# Source visualisation functions
source(here::here("src", "scenario_comparison_visualisation.R"))

# Check if results exist
if (!exists("scenario_comparison_results")) {
  stop("Scenario results not found. Please run the scenario_comparison_analysis chunk first.")
}

# Create comprehensive dashboard
dashboard <- create_scenario_comparison_dashboard(
  scenario_results = scenario_comparison_results,
  
  # USER CHOICE: Customise dashboard
  include_temperature = TRUE,
  include_emissions = TRUE,
  include_controls = TRUE,
  include_costs = TRUE,
  
  # USER CHOICE: Styling
  color_palette = "viridis",
  save_plot = TRUE,
  width = 297,      # A4 landscape width (mm)
  height = 210,     # A4 landscape height (mm)
  verbose = TRUE
)

# Display the dashboard
print(dashboard)
```

```{r parameter_importance_analysis, eval=FALSE}
# ==============================================================================
# Solution generation for multiple parameters & single SSP
# ==============================================================================
# Parameter importance analysis using Latin Hypercube Sampling (LHS) parameter sets
# This explores the solution space across multiple parameter combinations
# Provides output for sensitivity analysis

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "parameter_importance.R"))

# USER CHOICE: Modify these settings as needed
scenario_to_analyse <- "SSP4-Baseline"

# Run parameter importance analysis (saves automatically)
parameter_importance_results <- run_parameter_importance(
  parameter_df = parameter_df,
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenario = scenario_to_analyse,
  mitigation_delay_years = 0,
  cdr_delay_years = 0,
  use_parallel = TRUE,
  save_results = TRUE,
  verbose = FALSE
)

# Filter parameter importance results to remove problematic runs
# This removes runs that fail to return temperature to 1.5°C by 2100

# USER CHOICE: Set maximum acceptable final temperature (default: 1.6°C)
max_acceptable_temperature <- 1.6

# Apply filtering
parameter_importance_results_cleaned <- filter_importance_results(
  importance_results = parameter_importance_results,
  max_final_temperature = max_acceptable_temperature,
  verbose = TRUE
)
```

```{r parameter_importance_visualisation, eval=FALSE}
# ==============================================================================
# Plot multiple parameters & single SSP results
# ==============================================================================
# Create visualisations for parameter importance analysis results

# Source parameter importance visualisation functions
source(here::here("src", "parameter_importance_visualisation.R"))

# Check if results exist
if (!exists("parameter_importance_results_cleaned")) {
  stop("Parameter importance results not found. Please run the parameter_importance_analysis and parameter_importance_filtering chunks first.")
}

# Create and display dashboard with insights
parameter_importance_dashboard <- create_parameter_importance_dashboard(
  importance_results = parameter_importance_results_cleaned,
  save_plot = TRUE,
  print_insights = FALSE,
  verbose = FALSE
)

print(parameter_importance_dashboard)
```

```{r export_parameter_importance_for_sobol, eval=FALSE}
# ==============================================================================
# Export multiple SSPs & single parameter set results
# ==============================================================================
# Export parameter importance results for external Sobol indices analysis
# Creates CSV files with parameter values and output variables

# Source required functions
source(here::here("src", "parameter_importance.R"))

# Check if cleaned results exist
if (!exists("parameter_importance_results_cleaned")) {
  stop("Cleaned parameter importance results not found. Please run parameter_importance_filtering chunk first.")
}

# Export results for Sobol analysis (saves automatically)
export_for_sobol_analysis(
  importance_results = parameter_importance_results_cleaned,
  verbose = FALSE
)
```

```{r delayed_deployment_analysis, eval=FALSE}
# ==============================================================================
# Delayed Deployment Analysis
# ==============================================================================
# Analyse the effects of delaying mitigation and/or CDR deployment
# 
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: Scales with (max_delay / step_size)² × number of scenarios
#          Example: 50-year max, 10-year steps, 5 scenarios = ~30 minutes

# Source required functions
source(here::here("src", "optimal_control_core.R"))
source(here::here("src", "delayed_deployment.R"))

# USER CHOICE: Delay range and granularity
max_delay_years <- 70        # Maximum delay to test (years)
delay_step_size <- 1          # Step size (1 = yearly, 10 = every decade)

# USER CHOICE: Scenarios to analyse
# Options: "all" or vector like c("SSP2-Baseline", "SSP3-Baseline")
scenarios_to_analyse <- "all"

# Run delayed deployment analysis
delayed_deployment_results <- run_unified_delayed_deployment(
  parameter_df = parameter_df[1, ],      # Must be single row
  emissions_df = emissions_df,
  economic_df = economic_df,
  scenarios = scenarios_to_analyse,
  max_delay_years = max_delay_years,
  delay_step_size = delay_step_size,
  use_parallel = TRUE,
  save_results = TRUE,
  verbose = FALSE
)

# Display summary
cat("\n=== DELAYED DEPLOYMENT ANALYSIS COMPLETE ===\n")
cat("Delay range tested: 0 -", max_delay_years, "years\n")
cat("Total runs:", length(delayed_deployment_results$results), "\n")
```

```{r delayed_deployment_visualisation, eval=FALSE}
# ==============================================================================
# Delayed Deployment Visualisation
# ==============================================================================
# Create heatmap visualisations showing how deployment delays affect outcomes
# OUTPUTS: Saved automatically to figs/ directory

# Source visualisation functions
source(here::here("src", "delayed_deployment_visualisation.R"))

# Check if results exist
if (!exists("delayed_deployment_results")) {
  stop("Delayed deployment results not found. Please run the delayed_deployment_analysis chunk first.")
}

# Create abatement cost field (mitig cost + remov_cost)
# Calculate abatement cost
delayed_deployment_results$combined_results <- delayed_deployment_results$combined_results %>%
  mutate(abatement_cost = mitig_cost + remov_cost)

# Create main dashboard
dashboard <- create_delayed_deployment_dashboard(
  deployment_results = delayed_deployment_results,
  
  # USER CHOICE: Variables to plot
  # Options: "peak_temperature", "years_above_1p5", "abatement_cost", 
  #          "temp_cost", "total_cost", "mitig_cost", "remov_cost"
  variables = c("total_cost", "abatement_cost", "temp_cost"),
  
  # USER CHOICE: Visual features
  add_contours = TRUE,
  contour_alpha = 0.6,
  add_arrows = FALSE,
  show_infeasible = FALSE,
  
  # USER CHOICE: Styling
  color_palettes = NULL,
  shared_scale = TRUE,
  # Enable percentile capping
  use_scale_limits = TRUE,
  scale_limit_percentile = 70,
  save_plot = TRUE,
  width = 297,
  height = 210,
  verbose = FALSE
)

# Display the dashboard
print(dashboard)
```

```{r delayed_deployment_prebuilt_visualisation, eval=FALSE}
# ==============================================================================
# Convenience Wrapper Functions for Common Visualisations
# ==============================================================================
# These provide pre-configured shortcuts for standard analysis types

# Temperature and Overshoot Analysis
temperature_dashboard <- create_temperature_delay_dashboard(
  deployment_results = delayed_deployment_results,
  save_plot = FALSE,
  filename = NULL
)
print(temperature_dashboard)

# Cost Analysis with Gradient Arrows
cost_dashboard <- create_cost_delay_dashboard(
  deployment_results = delayed_deployment_results,
  cost_variable = "abatement_cost",  # Options: "abatement_cost", "temp_damage_cost", "total_cost"
  arrow_scale = 3.0,
  min_magnitude = 0,
  save_plot = FALSE,
  filename = NULL
)
print(cost_dashboard)

# Comprehensive Cost Comparison
comprehensive_cost_dashboard <- create_comprehensive_cost_delay_dashboard(
  deployment_results = delayed_deployment_results,
  shared_scale = TRUE,
  save_plot = FALSE,
  filename = NULL,
  height = 260
)
print(comprehensive_cost_dashboard)
```

```{r scenario_comparison_analysis_capacity, eval=FALSE}
# ==============================================================================
# Scenario Comparison Analysis with Capacity Constraints
# ==============================================================================
# Compare optimal control solutions across different SSP baseline scenarios
# with realistic deployment constraints on both mitigation and CDR.
#
# REQUIREMENTS: Must use single parameter set (parameter_df with 1 row)
# OUTPUTS: Saved automatically to output/ directory
# RUNTIME: ~5-15 minutes for 5 scenarios with parallel processing

# Source required functions
source(here::here("src", "optimal_control_capacity.R"))
source(here::here("src", "scenario_comparison_capacity.R"))
source(here::here("src", "capacity_helpers.R"))

# USER CHOICE: Select which SSP scenarios to compare
scenarios_to_compare <- c("SSP1-Baseline", "SSP2-Baseline", "SSP3-Baseline", 
                          "SSP4-Baseline", "SSP5-Baseline")

# USER CHOICE: Delay settings (typically keep at 0 for baseline comparison)
mitigation_delay <- 0  # Years before mitigation begins
cdr_delay        <- 0  # Years before CDR deployment begins

# ==============================================================================
# USER CHOICE: Mitigation Capacity Constraint
# ==============================================================================
# Enforces realistic ramp-up limits on mitigation deployment. This prevents
# unrealistic scenarios where mitigation jumps instantly to high levels.
# Examples: early coal/gas phase-out constraints, infrastructure buildout limits,
# political economy barriers to rapid emissions reductions.

use_mitigation_capacity_limit <- TRUE  # Set TRUE to activate

# Example 1: Exponential growth from current mitigation levels
# mitigation_capacity_function <- make_exponential_capacity(
#   u0 = 5,        # Start at 5 GtCO2/year (current global mitigation efforts)
#   gamma = 0.03,  # Grow at 3% per year
#   t_start = 2025
# )

# Example 2: Logistic S-curve (slow start, rapid middle, plateau)
#  mitigation_capacity_function <- make_logistic_capacity(
#   K = 60,        # Saturate at 40 GtCO2/year
#   r = 0.06,      # Steepness of transition
#   t_mid = 2045   # Midpoint of ramp-up
# )

# Example 3: Linear ramp with policy targets
# mitigation_capacity_function <- make_linear_capacity(
#   rho = 0.5,     # Add 0.5 GtCO2/year capacity each year
#   K = 50,        # Cap at 50 GtCO2/year
#   t_start = 2025
# )

# Example 4: Piecewise targets by decade (policy milestones)
# mitigation_capacity_function <- make_piecewise_capacity(
#   c("2025" = 5, "2030" = 10, "2040" = 20, "2050" = 35, "2070" = 45)
# )
 
# Moderate mitigation scaling
mitigation_capacity_function <- make_power_capacity(
  a = 0.0001,         # Moderate ambition
  alpha = 0.0001,   # Balanced learning/constraints
  t_start = 2020
)

# Mitigation scaling from current efforts
# mitigation_capacity_function <- make_logistic_from_zero(
#   g_initial = 5,     # Current ~5 GtCO2/year mitigation
#   K = 50,            # High deployment scenario
#   r = 0.03,          # Moderate growth
#   t_start = 2025
# )

# ==============================================================================
# USER CHOICE: CDR Capacity Constraint
# ==============================================================================
# Enforces realistic ramp-up limits on CDR deployment. This prevents unrealistic
# spikes (e.g., 40 GtCO2/year immediately). CDR must start from near-zero and
# grow based on technology learning, infrastructure, and investment.

use_cdr_capacity_limit <- TRUE  # Set TRUE to activate

# Example 1: Exponential growth from small base (realistic for new tech)
# cdr_capacity_function <- make_exponential_capacity(
#   u0 = 0.1,      # Start at 0.1 GtCO2/year (current CDR ~0.01)
#   gamma = 0.05,  # Grow at 5% per year
#   t_start = 2025
# )

# Example 2: Logistic S-curve (technology learning curve)
# cdr_capacity_function <- make_logistic_capacity(
#   K = 50,        # Saturate at 10 GtCO2/year (physical/economic limit)
#   r = 0.03,      # Steepness
#   t_mid = 2060   # Midpoint year
# )

# Example 3: Linear ramp (steady buildout)
# cdr_capacity_function <- make_linear_capacity(
#   rho = 0.2,     # Add 0.2 GtCO2/year capacity each year
#   K = 15,        # Cap at 15 GtCO2/year
#   t_start = 2030 # Start ramp in 2030
# )

# Example 4: Aggressive early deployment with plateau
# cdr_capacity_function <- make_linear_capacity(
#   rho = 0.5,     # Rapid initial buildout
#   K = 20,        # High saturation
#   t_start = 2025
# )

# Conservative CDR deployment
# cdr_capacity_function <- make_power_capacity(
#   a = 6,       # Low deployment ambition
#   alpha = 0.35,  # Strong diminishing returns
#   t_start = 2020
# )

# CDR with three-phase deployment
cdr_capacity_function <- make_logistic_from_zero(
  g_initial = 1,  # Current CDR levels
  K = 100,            # Maximum realistic capacity
  r = 0.05,          # Moderate acceleration
  t_start = 2025
)

# ==============================================================================
# Run scenario comparison with capacity constraints
# ==============================================================================
# Both mitigation and CDR constraints can be active simultaneously.
# The solver will optimize within the feasible deployment space defined
# by the capacity functions.

scenario_comparison_results <- run_scenario_comparison(
  parameter_df                   = parameter_df[1, ],    # Must be single row
  emissions_df                   = emissions_df,
  economic_df                    = economic_df,
  scenarios                      = scenarios_to_compare,
  mitigation_delay_years         = mitigation_delay,
  cdr_delay_years                = cdr_delay,
  use_mitigation_capacity_limit  = use_mitigation_capacity_limit,  # Mitigation constraint
  mitigation_capacity_function   = mitigation_capacity_function,    # Mitigation capacity curve
  use_cdr_capacity_limit         = use_cdr_capacity_limit,          # CDR constraint
  cdr_capacity_function          = cdr_capacity_function,           # CDR capacity curve
  use_parallel                   = TRUE,                            # Parallel processing
  save_results                   = TRUE,                            # Save to output/
  verbose                        = FALSE                            # Progress reporting
)

# Display summary
if (!is.null(scenario_comparison_results$comparison_summary)) {
  print(scenario_comparison_results$comparison_summary)
}

# ==============================================================================
# Interpretation Notes
# ==============================================================================
# 
# WHEN 1.5°C IS IMPOSSIBLE:
# If capacity constraints make returning to 1.5°C impossible, the solver returns
# the "best" result it can find - the solution that gets closest to the target
# cumulative emissions. Check these fields to assess the result:
#
#   - final_temperature: Actual temperature achieved (may exceed 1.5°C)
#   - emission_gap: How close to target emissions (GtCO2)
#   - temperature_anomaly: Full time series showing overshoot trajectory
#
# CAPACITY BINDING INDICATORS:
# To see if capacity constraints are actively limiting deployment:
#
#   - Plot qty_mitig vs mitigation_capacity_function(years)
#   - Plot qty_remov vs cdr_capacity_function(years)
#   - If they track closely, the constraint is binding
#
# OVERSHOOT DYNAMICS:
# Capacity constraints typically force temporary temperature overshoot because
# deployment can't scale fast enough to avoid it. The solver minimizes total
# cost subject to the deployment constraints, which may mean accepting overshoot.
#
```

# ==============================================================================
# Workflow Complete
# ==============================================================================
# 
# If all chunks have run successfully, you should now have:
# ✓ Processed emissions and economic data for all SSP scenarios
# ✓ Generated parameter sets (either fixed or LHS)
# ✓ Optimal control solutions for selected scenarios
# ✓ Publication-ready visualisations in figs/ directory
# ✓ Results files saved in output/ directory
#
# NEXT STEPS:
# - Load saved RDS files for detailed analysis
# - Customise visualisations by modifying plotting parameters
# - Run different parameter sets or scenarios
# - See USER_GUIDE.md for advanced usage and troubleshooting
#
# GETTING HELP:
# - Email: nina.rynne@griffithuni.edu.au
# - Repository: https://github.com/nina-rynne/MACROM
# - Issues: Submit via GitHub issue tracker
#
# Thank you for using MACROM!
# ==============================================================================